---
title: "Reticulate"
author: "Kevin O'Brien"
date: "5 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(reticulate)
```

## ``repl_python``: Run a Python REPL
#### Description 
This function provides a Python REPL in the R session, which can be used to interactively run
Python code. All code executed within the REPL is run within the Python main module, and any
generated Python objects will persist in the Python session after the REPL is detached.
#### Usage
<pre><code>
repl_python(module = NULL, quiet = getOption("reticulate.repl.quiet",
default = FALSE))
</code></pre>

#### Arguments
* module An (optional) Python module to be imported before the REPL is launched.
* quiet Boolean; print a startup banner when launching the REPL? If TRUE, the banner will be suppressed.

#### Details
When working with R and Python scripts interactively, one can activate the Python REPL with
repl_python(), run Python code, and later run exit to return to the R console.


#### Examples

```{r}
## Not run:
# enter the Python REPL, create a dictionary, and exit
repl_python()
dictionary = {'alpha': 1, 'beta': 2}
exit
# access the created dictionary from R
py$dictionary
# $alpha
# [1] 1
#
# $beta
# [1] 2
## End(Not run)
```

See Also
py, for accessing objects created using the Python REPL.