---
title: "escape Escape/quote a string."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dbplyr)
```


#### Description 
Escape/quote a string.

#### Usage
<pre><code>
escape(x, parens = NA, collapse = " ", con = NULL)
sql_vector(x, parens = NA, collapse = " ", con = NULL)
</code></pre>
  
#### Arguments
* x An object to escape. Existing sql vectors will be left as is, character vectors are escaped with single quotes, numeric 
 vectors have trailing .0 added if they’re whole numbers, identiﬁers are escaped with double quotes. 
* parens, collapse Controls behaviour when multiple values are supplied. 
  - parens should be a logical ﬂag, or if NA, will wrap in parens if length > 1. Default behaviour: lists are always wrapped in parens and separated by commas,identiﬁers are separated by commasand never wrapped,atomic vectors are separated by spaces and wrapped in parens if needed. 
 * con Database connection. If not speciﬁed, uses SQL 92 conventions.

#### Examples
```{r}
# Doubles vs. integers 
escape(1:5) 
escape(c(1, 5.4))
# String vs known sql vs. sql identifier 
escape("X") 
escape(sql("X")) 
escape(ident("X"))
# Escaping is idempotent 
escape("X") 
escape(escape("X")) 
escape(escape(escape("X")))
```
