---
title: " ``remote_name`` Metadata about a remote table"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dbplyr)
```


#### Description 

* ``remote_name()`` gives the name remote table, or NULL if it’s a query. 
* ``remote_query()`` gives the text of the query,and ``remote_query_plan()`` the query plan ( as computed by the remote database). 
* ``remote_src()`` and ``remote_con()`` give the dplyr source and DBI connection respectively.

### Usage
<pre><code>
remote_name(x)
remote_src(x)
remote_con(x)
remote_query(x)
remote_query_plan(x)
</code></pre>

#### Arguments 
* ``x``: Remote table, currently must be a tbl_sql.
* ``Value``: The value, or NULL if not remote table, or not applicable. For example, computed queries do not have a "name"

#### Examples

```{r}
mf <- memdb_frame(x = 1:5, y = 5:1, .name = "blorp") 
remote_name(mf) 
remote_src(mf) 
remote_con(mf) 
remote_query(mf)
mf2 <- dplyr::filter(mf, x > 3) 
remote_name(mf2) 
remote_src(mf2) 
remote_con(mf2) 
remote_query(mf2)

```
