translate_sql Translate an expression to sql.

#### Description 

Translate an expression to sql.

#### Usage
<pre><code>
translate_sql(..., con = NULL, vars = character(), vars_group = NULL, vars_order = NULL, vars_frame = NULL, window = TRUE)
translate_sql_(dots, con = NULL, vars_group = NULL, vars_order = NULL, vars_frame = NULL, window = TRUE, context = list())
</code><pre>

#### Arguments 
* ..., dots Expressions to translate. translate_sql() automatically quotes them for you. translate_sql_() expects a list of already quoted objects. con An optional database connection to control the details of the translation. The default, NULL, generates ANSI SQL. vars Deprecated. Now call partial_eval() directly. vars_group, vars_order, vars_frame Parameters used in the OVER expression of windowed functions. window Use FALSE to suppress generation of the OVER statement used for window functions. This is necessary when generating SQL for a grouped summary. context Use to carry information for special translation cases. For example, MS SQL needs a different conversion for is.na() in WHERE vs. SELECT clauses. Expects a list.
* Basetranslation The base translator, base_sql, provides custom mappings for ! (to NOT), && and & to AND, || and | to OR, ^ to POWER, %>% to %, ceiling to CEIL, mean to AVG, var to VARIANCE, tolower to LOWER, toupper to UPPER and nchar to LENGTH. c() and : keep their usual R behaviour so you can easily create vectors that are passed to sql.

All other functions will be preserved as is. R’s inﬁx functions (e.g. %like%) will be converted to their SQL equivalents (e.g. LIKE). You can use this to access SQL string concatenation: || is mapped to OR, but %||% is mapped to ||. To suppress this behaviour, and force errors immediately whendplyrdoesn’tknowhowtotranslateafunctionitencounters,usingsetthedplyr.strict_sql option to TRUE. You can also use sql() to insert a raw sql string.
* SQLitetranslation The SQLite variant currently only adds one additional function: a mapping from sd() to the SQL aggregation function STDEV.

#### Examples
```{r}
# Regular maths is translated in a very straightforward way 
translate_sql(x + 1) 
translate_sql(sin(x) + tan(y))
```
```{r}
# Note that all variable names are escaped translate_sql(like == "x") 
# In ANSI SQL: "" quotes variable _names_, '' quotes strings
# Logical operators are converted to their sql equivalents 
translate_sql(x < 5 & !(y >= 5)) 
```
```{r}
# xor() doesn't have a direct SQL equivalent translate_sql(xor(x, y))
# If is translated into case when translate_sql(if (x > 5) "big" else "small")
# Infix functions are passed onto SQL with % removed translate_sql(first %like% "Had%") 
translate_sql(first %is% NULL) 
translate_sql(first %in% c("John", "Roger", "Robert"))
```
```{r}
# And be careful if you really want integers 
translate_sql(x == 1) translate_sql(x == 1L)
# If you have an already quoted object, use translate_sql_: x <- quote(y + 1 / sin(t)) 
translate_sql_(list(x))
```
```{r}
# Windowed translation -------------------------------------------# 
Known window functions automatically get OVER() translate_sql(mpg > mean(mpg))
# Suppress this with window = FALSE translate_sql(mpg > mean(mpg), window = FALSE)
# vars_group controls partition: 
translate_sql(mpg > mean(mpg), vars_group = "cyl")
window_order 17
```
```{r}
# and vars_order controls ordering for those functions that need it 
translate_sql(cumsum(mpg)) 
translate_sql(cumsum(mpg), vars_order = "mpg")
```
